---
title: "Review Wind Exposure"
author: "Maite"
date: "3 11 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Review: Wind exposure and LST 
```{r, message=F, warning=F,echo=F}
library(raster)
library(mapview)
library(rasterVis)
library(ncdf4) # package for netcdf manipulation
library(rgdal)
library(corrplot)


aoipath <-  "E:/new_downscaling/aoi/" # this must point to an existing directory with "Levy_MDV_actually.shp"
aoip <- list.files(aoipath, pattern="actually.shp", full.names = T)
aoi <- readOGR(aoip)
antaproj <- crs("+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +k=1 +x_0=0 +y_0=0 +datum=WGS84 +units=m
+no_defs +ellps=WGS84 +towgs84=0,0,0")
aoianta <- spTransform(aoi, antaproj)
aoiwgs <- spTransform(aoianta, crs("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs "))


```


http://www.brianbuma.com/news/2017/12/29/quantifying-wind-exposure-in-r

First, the following code quickly calculates - based on user supplied wind direction, deflection angles, and search distances (how far upwind a barrier should matter) - relative exposure:

```{r, eval=F, include=F}
# function

windout.iter <- function(dem, deflect, angles, max.dist) {
    #note that the dem raster must be in planar coordinates
    
    #for smaller datasets:
    #dem <- readAll(dem)     #if in memory
    res <- res(dem)
    
    # do not ignore the case where x and y resolution are not equal
    stopifnot(all.equal(res[1], res[2]))
    xr <- res[1]
    
    #number of distances to check, basically goes every other cell for speed.
    num.dist <- round(max.dist / xr / 2)
    distance <- seq(xr, max.dist, length.out=num.dist)
    result <- list()
    j <- 1
    
    for (d in deflect) {
        midrow <- cellFromRow(dem,rownr=1)    #note this does the top one
        elev <- extract(dem,midrow)
        coords <- xyFromCell(dem, midrow)
    
        radangle <- (angles+90) * pi/180  #convert to radians.
        dcosangle <- -cos(radangle) * distance
        dsinangle <- sin(radangle) * distance
        x <- apply(coords[,1,drop=FALSE], 1, function(j) j + dcosangle)
        y <- apply(coords[,2,drop=FALSE], 1, function(j) j + dsinangle)
        xy <- cbind(as.vector(x), as.vector(y))
    
        comp.elev <- extract(dem, xy)
        comp.elev <- matrix(comp.elev, ncol=num.dist, byrow=TRUE)
        comp.elev <- comp.elev - elev
        comp.elev <- t(t(comp.elev) / distance)
        #notAllNA <- rowSums(is.na(comp.elev)) != num.dist
        ang <- atan(comp.elev) * (180 / pi)
    
        r <- apply(ang,1,max)
    
        r <- r<=d
        result[[j]] <- r*1
        j <- j+1
        }
    
    output <-simplify2array(result)
    output <- apply(output,1,sum)
    output <- output+1
    outputs <- list()
    outputs[[1]] <- output
    outputs[[2]] <- coords
    return(outputs)
}

```


This next code then loops through a DEM, first subsetting out an area the size of the max distance (that keeps things fast) and then calculating all wind directions desired and averaging the results.  To do this in practice, one needs to know the distribution of storm-force wind directions, and builds the directions based off that (to weight a given direction, one could either calculate twice or simply duplicate).  The following works at a 5km distance, several deflection angles, and is oriented around S, SE, and SW wind:

```{r, eval=F, include=F}

#load single big DEM
dem <- raster(paste0(dempath, "DEM_30m_", areaname,"_clean_aoi_filled_mask_new.tif"))

plot(dem,maxpixels=60000)
e <- extent(c(399967.3, 409091, -1287161, -1279727 ))
dem <- crop(dem,e)   #If you want to do a focal area

# #set projection
# proj.def <- "+proj=utm +ellps=WGS84 +zone=8 +units=m"  #Adjust as needed.
# dem <- projectRaster(dem, crs=proj.def)

storage <- matrix(nrow=nrow(dem),ncol=ncol(dem))

#set parameters
max.dist <- 5000
deflect <- c(1,3,5,7,9,11,13,14)
angles <- c(135,180,225)

iter <- 1:nrow(dem)    #to 3772 now
r <- res(dem)[1]


for (i in iter) {
    temp.extent <- extent(dem)
    temp.extent@ymax <- temp.extent@ymax-(i*r)    #*res(dem)[1] to avoid top
    temp.extent@ymin <- temp.extent@ymax-(i*r+max.dist+r)

    temp.dem <- crop(dem,temp.extent)
    temp1 <- windout.iter(temp.dem,deflect,angles[1],max.dist)[[1]]
    temp2 <- windout.iter(temp.dem,deflect,angles[2],max.dist)[[1]]
    temp3 <- windout.iter(temp.dem,deflect,angles[3],max.dist)[[1]]

    #temp.coords <- SpatialPoints(temp.coords)
    temp <- apply(cbind(temp1,temp2,temp3),1,mean)
    #t.loc <- cellFromXY(storage,temp.coords)

    storage[i,] <- temp    
    print(i)
    removeTmpFiles(h=0)   #This is needed large processing jobs, which crash otherwise.
    rm(temp.dem)
    gc()
}

gc()
t <- dem   #this creates a place to put the calculated values
t[] <- storage

mapview(t)+
  mapview(dem)

writeRaster(t, "D:/LST_paper_review/exposure/exp_example.tif",
            format="GTiff",overwrite=T)


```


## Get AMPS data

Metadata: 

 float U[west_east_stag,south_north,bottom_top,Time]   (Chunking: [193,225,15,1])  (Compression: shuffle,level 2)
      FieldType: 104
      MemoryOrder: XYZ
      description: x-wind component
      units: m s-1
      stagger: X
      coordinates: XLONG_U XLAT_U XTIME
  float V[west_east,south_north_stag,bottom_top,Time]   (Chunking: [192,226,15,1])  (Compression: shuffle,level 2)
      FieldType: 104
      MemoryOrder: XYZ
      description: y-wind component
      units: m s-1
      stagger: Y
      coordinates: XLONG_V XLAT_V XTIME
  float W[west_east,south_north,bottom_top_stag,Time]   (Chunking: [192,225,16,1])  (Compression: shuffle,level 2)
      FieldType: 104
      MemoryOrder: XYZ
      description: z-wind component
      units: m s-1
      stagger: Z
      coordinates: XLONG XLAT XTIME


Info on U and V for wind components http://colaweb.gmu.edu/dev/clim301/lectures/wind/wind-uv 

third dimension = height

### read stuff from netcdf
```{r, echo=F, message=F, warning=F}
# from library(Rsenal)
uv2wdws <- function(u,v) {
  degrees <- function(radians) 180 * radians / pi
  mathdegs <- degrees(atan2(v, u))
  wdcalc <- ifelse(mathdegs > 0, mathdegs, mathdegs + 360)
  wd <- ifelse(wdcalc < 270, 270 - wdcalc, 270 - wdcalc + 360)
  ws <- sqrt(u^2 + v^2)
  return(cbind(wd, ws))
}



ad <- "D:/LST_paper_review/AMPS_data/"
adf <- list.files(ad, full.names=T, pattern = "14.nc$")

for(i in seq(adf)){
    nc_data <- nc_open(adf[i])
  # Save the print(nc) dump to a text file
  {
      sink(paste0(tools::file_path_sans_ext(adf[i]), '_metadata.txt'))
   print(nc_data)
      sink()
  }
}
  
windX <- ncvar_get(nc_data, "U")
windY <- ncvar_get(nc_data, "V")
windZ <- ncvar_get(nc_data, "W")

lon <- ncvar_get(nc_data, "XLONG")
lat <- ncvar_get(nc_data, "XLAT")

nc_close(nc_data) 

wind_l <- list(windX, windY, windZ)
names(wind_l) <- c("U", "V", "Z")


par(mfrow=c(1,3))
hist(windX)
hist(windY)
hist(windZ)
```


```{r, message=F, warning=F}
dim(lon)
dim(lat)

dim(windX) 
dim(windY)
dim(windZ)
```



The coordinates seen in NCview 

Not sure how to get them, really doesn't make much sense to me from the metadata and ncdf file... 

They are: 

```{r, message=F, warning=F,echo=F}
extent_NCView <- extent(c(152.695, 180, -81.839, -74.7438))
ep <- as(extent_NCView, 'SpatialPolygons')
crs(ep) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0")
mapview(ep)

```


### Look at the necessary flips and t for coordinate rasters
```{r, message=F, warning=F,echo=F}

par(mfrow=c(1,2))
rlon <- raster(lon)
plot(rlon, main="lon original position")

rlon <- raster(t(lon))
rlon <- flip(rlon, direction='y')
plot(rlon, main="lon transposed and flipped")


```

This shows, that the matrix is transposed, we also need to flip, as the actual data situation proved. 

The lat data: 

```{r, message=F, warning=F,echo=F}
rlat <- raster(t(lat))
rlat <- flip(rlat, direction="y") 
par(mfrow=c(1,1))
plot(rlat, main="lat transposed and flipped")
```


### deal with the 180° problem

Now, for the untransposed and unflipped data, we need to cut to the extent witnessed in NCView.

For lon max, we cut to 180° by taking all columns, that are positive, i.e. not using the -177 etc. 

```{r,echo=TRUE}
extent_NCView

rlon <- raster(lon)

dim(rlon)
dim(lon)
```

dim(lonnew)

```{r, message=F, warning=F,echo=F}
# Cut xmax (lon max) to 180
# in the column of lon, where do values get negative for the first time
cutoff_index <- which(lon[,1]<1)[1]
lonnew <- lon[1:(cutoff_index-1),]
dim(lonnew)
par(mfrow=c(1,2))
plot(raster(lonnew), main="new longitude")

# if we then t and flip
lonnew_ft <- raster(t(lonnew))
lonnew_ft <- flip(lonnew_ft)
plot(lonnew_ft, main="new longitude t and flipped")
```



# cropping all wind rasters to the new dimensions

On untransposed and unflipped data, we should be able to do the same as in lon.
```{r, message=F, warning=F,echo=F}
wind_l_cropped <- lapply(seq(wind_l), function(i){
  cat("org: ", dim(wind_l[[i]]))
  
  x <- wind_l[[i]][(1:(cutoff_index-1)),,]

  cat("new:", dim(x))
  x
  })
```



```{r, message=F, warning=F,echo=F}
# e <- c(150, 170, -79, -75)
wind_grid <- lapply(seq(wind_l), function(i){
  
  r_t1 <- wind_l_cropped[[i]][,,1] # get first height 
  rx <- raster(t(r_t1), # was t(windX_time1) looks better not transposed
            xmn=152.695, 
            xmx=max(lonnew), 
            ymn=-81.839, 
            ymx=-74.7438, 
            crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
  
  rx <- flip(rx, direction='y')
  
  rx <- crop(rx, aoiwgs)
  rx <- mask(rx, aoiwgs)
  # rxproj <- projectRaster(rx, crs=3031)
  # rxproj_aoi <- crop(rxproj, aoianta)
  # 
  # rxproj_aoi <- mask(rxproj, aoianta)

  
  # rx <- crop(rx, e)
  return(rx)
})


names(wind_grid) <- c("U", "V", "Z")

mapview(wind_grid$U,map.types="Esri.WorldImagery",na.color="#00000000")+
  mapview(wind_grid$V, na.color="#00000000")+mapview(aoianta, alpha.regions=0, lwd=4)
```

## wind direction and speed 

```{r,message=F, warning=F,echo=F}
convert_to_wd_and_ws <- function(wind_l, height=1){
  wd_ws <- uv2wdws(as.vector(wind_l$U[,,height]), 
          as.vector(wind_l$V[,,height]))
  
  wd_ws <- uv2wdws(wind_grid$U[],
          wind_grid$V[])
  
  dim(wd_ws)
  
  wd <- wind_grid$V
  ws <- wind_grid$V
  
  wd[] <- wd_ws[,1]
  ws[] <- wd_ws[,2]
  
  # 
  # wd <- crop(wd, aoiwgs)
  # ws <- crop(ws, aoiwgs)
  
  return(list(wd = wd, 
              ws = ws))
}

wd_ws_aoi <- convert_to_wd_and_ws(wind_l = wind_grid, height=1)

mapview(wd_ws_aoi$wd,map.types="Esri.WorldImagery",na.color="#00000000")+
  mapview(wd_ws_aoi$ws,map.types="Esri.WorldImagery",na.color="#00000000")+
  mapview(aoianta, alpha.regions=0, lwd=4)
```



```{r, message=F, warning=F, echo=F, eval=F, include=F}

windField <- stack(wd_ws_aoi$ws, wd_ws_aoi$wd)
names(windField) <- c('magnitude', 'direction')


# from here https://oscarperpinan.github.io/bookvis/spatial.html#vector
vectorTheme <- BTCTheme(regions = list(alpha = 0.7))

vectorplot(windField, isField = TRUE,
           aspX = 5, aspY = 5,
           narrows=900,lwd.arrows=2,
           length=unit(0.03, 'npc'),
           scaleSlope = FALSE, 
           key.arrow = list(label = 'm/s'),
           par.settings = vectorTheme,
           colorkey = FALSE,
           scales = list(draw = FALSE))

# myTheme <- streamTheme(
#     region = rev(brewer.pal(n = 4, "Greys")),
#     symbol = rev(brewer.pal(n = 9, "Blues")))
# 
# streamplot(windField, isField = TRUE,
#            par.settings = myTheme,
#            droplet = list(pc = 12),
#            streamlet = list(L = 5, h = 5),
#            scales = list(draw = FALSE),
#            panel = panel.levelplot.raster)

vectorplot(wd_ws_aoi$ws, par.settings=BuRdTheme(),
           narrows=900,lwd.arrows=2,
           length=unit(0.02, 'npc'),
           key.arrow = list(label = 'm/s'))
```


```{r, eval=F, include=F}

wdws <- data.frame(wd=wd_ws_aoi$wd[], 
                   ws=wd_ws_aoi$ws[],
                   coordinates(wd_ws_aoi$ws))

ggplot(wdws, 
       aes(x = x , 
           y = y, 
           fill = ws, 
           angle = wd, 
           radius = scales::rescale(ws, c(.01, .1)))) +
    geom_raster() +
    geom_spoke(arrow = arrow(length = unit(.05, 'inches'))) + 
    scale_fill_distiller(palette = "RdYlGn") + 
    coord_equal(expand = 0) + 
    theme(legend.position = 'bottom', 
          legend.direction = 'horizontal')

```



## correlate wind exposure and LST
```{r, echo=F, message=F, warning=F}
expDir <- "D:/LST_paper_review/exposure/"
dem <- raster("C:/Users/mleza/OneDrive/Documents/PhD/work_packages/SurfaceMoisture/predictors/DEM_8m_MDV_filled_aoi.tif")
full_grid <- readRDS("C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/downscaleLST.MDV/data/full_size_grids_all_layers.RDS")
comp_ex2 <- readRDS("C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/downscaleLST.MDV/data/comp_ex2.RDS")
ex_2_all_layers <- readRDS("C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/downscaleLST.MDV/data/ex_2_all_layers.RDS")
full_dsc <- raster("D:/downscaling_after_talk/predictions/pred_rf_MYD11_L2.A2018316.1350.006.2bilinmod.tif")
exp150 <- raster(paste0(expDir, "Wind_Expo_150.tif"))

wind_eff <- raster(paste0(expDir, "Wind_Effect_900_full.tif"))
wind_eff <- mask(wind_eff, aoianta)

wind_exp <- raster(paste0(expDir, "Wind_Expo_900_full.tif"))
wind_exp <- mask(wind_exp, aoianta)

```


```{r, echo=F, message=F, warning=F, eval=F}

demex2 <- crop(dem, extent(comp_ex2))
writeRaster(demex2, paste0(expDir, "small_DEM_ex2.tif"))

writeRaster(full_grid$dem, paste0(expDir, "dem_30_full.tif"))

wind_exp_full <- raster(paste0(expDir, "Wind_Expo_900_full.tif"))
mapview(wind_exp_full)+mapview(full_grid$dem)

#### bring windspeed and winddirection into same format as full_grid

wdwss <- stack(wd_ws_aoi$wd, wd_ws_aoi$ws)
wdwss_p <- projectRaster(wdwss, full_grid)
wdwss_30 <- resample(wdwss_p, full_grid$dem)
names(wdwss_30) <- c("wd","ws")

writeRaster(wdwss_30$wd, paste0(expDir, "wd_full_30.tif"))
writeRaster(wdwss_30$ws, paste0(expDir, "ws_full_30.tif"))

# windspeed and winddirection for example 2
wd_proj <- projectRaster(wd_ws_aoi$wd, comp_ex2)
wd_proj_ex2 <- crop(wd_proj,comp_ex2)
wd_proj_ex2 <- resample(wd_proj_ex2, comp_ex2)

names(wd_proj_ex2) <- "wd_ex2"
comp_ex2 <- stack(comp_ex2, wd_proj_ex2)

writeRaster(wd_proj_ex2, paste0(expDir, "wd_ex2.tif"),
            overwrite=T)


```


I'll take the difference by: diff=downscaled-Landsat
That way, the values that are negative are where an underestimation happened in downscaling and 
positive values are areas where overestimation happened. 
```{r, echo=F, message=F, warning=F}
diff_full <- full_dsc-full_grid$L
mapview(diff_full)

diff <- comp_ex2$downscaled-comp_ex2$L
mapview(diff)
```

Overestimation happens mostly on the mountain peaks. 

```{r, echo=F, message=F, warning=F}

mapview(diff_full)+
  mapview(ex_2_all_layers$landcoverres)+
  mapview(exp150)

```


There does not seem to be such an explicit relation between LST and windspeed, 
but in fact windward position would be more interesting, i.e. wind direction and 
terrain 
```{r, echo=F, message=F, warning=F}
mapview(wd_ws_aoi$ws)+
  mapview(full_grid$L)


```

```{r, eval=F, include=F}


library("RSAGA", quietly = TRUE)
# rsaga.env()[1:5]
# rsaga.get.libraries()[1:60]
# rsaga.get.modules("ta_morphometry")

# made the saga grid dem by reading into SAGA GUi and writing from there 
rsaga.geoprocessor(lib = "ta_morphometry", module = "Wind Exposition Index",
                  param = list(DEM = paste0(expDir, "small_DEM_ex2.sgrd"),
                    #DEM = "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/SurfaceMoisture/predictors/DEM_8m_MDV_filled_aoi.sgrd", 
                               EXPOSITION = "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/SurfaceMoisture/predictors/Wind_Expo_900m.sgrd",
                  MAXDIST=900))

```


### correlate topographic wind exposition and windward / leeward index for 
example 2 and correlate with other variables
```{r, echo=F, warning=F, message=F}
# exp900 <- raster(paste0(expDir, "Wind_Expo_900.tif"))
# mapview(exp150)+mapview(exp900) # turned out to be the same

exp150proj <- projectRaster(exp150, diff)

wind_eff_ex2 <- crop(wind_eff, exp150)

s <- stack(exp150proj, diff, wind_eff_ex2)
names(s) <- c("wind_exp", "diff_L_dsc", "wind_effect")
sdf <- data.frame(s[])
cor.test(sdf$wind_exp, sdf$diff_L_dsc,
         use="pairwise.complete.obs")


mapview(exp150)+mapview(diff)+mapview(wind_eff_ex2)
par(mfrow=c(1,3))
plot(diff, main="predicted -observed LST")
plot(exp150, main="topographic wind exposure")
plot(wind_eff_ex2, main="wind effect AMPS")
```


```{r, echo=F, message=F, warning=F}
names(s) <- c("top_wind_exp", "diff_dsc_L", "wind_effect")
s_full <- stack(ex_2_all_layers, s)
s_full_df <- data.frame(s_full[])
names(s_full_df)

s_df_cor <- s_full_df[c("dem", "slope", "aspect", "soilraster", "diff_dsc_L", "top_wind_exp", "wind_effect","L")]
cor_mat_ex2 <- cor(s_df_cor, use="pairwise.complete.obs")

```



### correlate topographic wind exposition and windward / leeward index for 
full scene and correlate with other variables

```{r, echo=F, warning=F, message=F}
names(diff_full) <- "diff_dsc_L"
names(wind_eff) <- "wind_effect"
names(wind_exp) <- "top_wind_exp"
full_grid_wind <- stack(full_grid, wind_eff, wind_exp, diff_full)
s_full_stack <- full_grid_wind[[c("dem", "slope", "aspect", "soilraster", "diff_dsc_L", "top_wind_exp", "wind_effect","L")]]
s_full_df_cor <- data.frame(s_full_stack[])
cor_mat <- cor(s_full_df_cor, use="pairwise.complete.obs")
```

### Example 2
```{r, echo=F, warning=F, message=F, out.width="70%"}
par(mfrow=c(1,1), oma=c(0,0,2,0))
corrplot(cor_mat_ex2, method="number", 
         type="upper", tl.col="black",
         diag=F, insig="blank")
```

### Full Scene
```{r, echo=F, warning=F, message=F, out.width="70%"}
corrplot(cor_mat, method="number", 
         type="upper", tl.col="black",
         diag=F, insig="blank")
```

Example 2 (left) and full scene (right): 

Example2:
The wind exposition has a medium strong negative correlation with the 30m LST (L). 
The difference between observed and predicted LST has to do a lot with the terrain, 
the slope and the wind exposure. So it seems to actually be a very good predictor. 

Full Scene: 

## where does over/underestimation happen with respect to the terrain

```{r}
rs <- full_grid_wind[[c("dem", "L", "wind_effect", "diff_dsc_L")]]

rsdf <- data.frame(rs[])
corm <- cor(rsdf, use="pairwise.complete.obs")



mapview(full_grid_wind$dem)+
  mapview(full_grid_wind$L)+
  mapview(full_grid_wind$wind_effect)+
  mapview(full_grid_wind$diff_dsc_L)
```

Overestimation mostly on the top of the mountains. 


```{r}
full_grid$landcoverres
writeRaster(full_grid$landcoverres, paste0(dempath, "landcoverres.tif"))

```

